---
title: "Juvenile Snowshoe Hares: Exploratory Report"
author: "Katelyn Toigo"
date: "11/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(kableExtra)
library(ggbeeswarm)
library(effsize)
library(broom)
```

## Introduction

This report looks at juvenile snowshoe hares on the Bonanza Creek Long Term Ecological Research (LTER) site between 1998 and 2012 to better understand how many were caught, the relationship between sex and weight, and the relationship between weight and hind foot length.

## Data and Analyses

Snowshoe hare physical data were collected made available by Dr. Knut Kielland et al. at the Bonanza Creek Long Term Ecological Research (LTER) site in Alaska. The data has 3197 observations made between 1998-2012. Following exploratory data visualization, weights (g) between male and female hares are compared by two-sample t-tests using a significance level ($\alpha$) of 0.05 throughout. Differences between groups are described by Cohen's *d* effect size. The relationship between weight (g) and hind foot length (mm) is explored by simple linear regression. All analyses are in R version 4.0.2 using RStudio version 1.3.1056.

## Exploratory findings
### Annual juvenile hare trap counts

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the data
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Create a new subset of the data containing only grid, date, and age. Age is filtered to only juveniles.  A new date column is created and the year is selected and made its own column. The data is grouped by year and the number of juveniles caught per year are counted. Count() combines group_by + summarize + n.
juv_year <- bonanza_hares %>% 
  filter(age == "j") %>% 
  select(grid, date, age) %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  group_by(year) %>% 
  count(age == "j")
  
# Ggplot of the dataset is created.
ggplot(data = juv_year, aes(x=year, y=n)) +
  geom_bar(stat="identity") +
  labs(title = "Juvenile Hares Caught by Year", x = "Year", y = "Number of Juveniles")
```

**Figure 1.** The number of juvenile hares captured each year at Bonanza Black Spruce, Bonanza Mature, and Bonanza Riparian. Data: Kielland, Chapin, Ruess (2017).

The maximum number of juvenile hares were captured in 1999, and after this the number of juveniles captured decreased every year until 2004. The minimum number caught were 0 in 2002 and 2009. The number captured tended to fluctuate.  The number captured each year also depends on the amount of sampling. If sampling cannot be consistent, it might be better to compare data that is standardized, such as number of juveniles caught by day.


### Visualize juvenile hare weights

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# A new subset of data is created
weights_sex_viz <- bonanza_hares %>% 
  select(grid, sex, weight, age) %>% 
  filter(age == "j")

#The following is used to rename values within the table. The dollar sign tells it to look in a specific column.
weights_sex_viz$grid[weights_sex_viz$grid == "bonbs"] <- "Bonanza Black Spruce"
weights_sex_viz$grid[weights_sex_viz$grid == "bonmat"] <- "Bonanza Mature"
weights_sex_viz$grid[weights_sex_viz$grid == "bonrip"] <- "Bonanza Riparian"

weights_sex_viz$sex[weights_sex_viz$sex == "f"] <- "Female"
weights_sex_viz$sex[weights_sex_viz$sex == "m"] <- "Male"
```


```{r, include = FALSE}
# This code is being run so that I can add mean points to the beeswarm graph.
juv_weights_summary <- weights_sex_viz %>% 
  group_by(sex) %>% 
  summarize(
    mean_weight = round(mean(weight, na.rm = TRUE),2),
    sd_weight = round(sd(weight, na.rm = TRUE),2),
    n = n())
```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Beeswarm graph. Geom_beeswarm was used to plot the points and boxplot was used to plot the boxes.  Facet_wrap to separate into three separate graphs by the grid (site).  Labs is used to change the labels.
ggplot(data = weights_sex_viz, aes(x=sex, y=weight)) + 
  geom_beeswarm(aes(color = sex))+
  facet_wrap(~grid)+
  geom_boxplot(fill = NA, width = 0.3, outlier.color = NA) +
  labs(title = "Sex vs. Weight Comparison for Juvenile Snowshoe Hares", x="Sex", y="Weight (grams)")
```

**Figure 2.** Sex versus weight (grams) comparisons for juvenile snowshoe hares at three sites: Bonanza Black Spruce, Bonanza Mature, and Bonanza Riparian. Red (female), teal (male) and grey (NA - unidentified as male or female) points indicate individual observations for weight (grams) of juvenile snowshoe hares. Box endpoints indicate the 25^th^ and 75^th^ percentile values; the black line inside each box indicates the median value for each sex. Data: Kielland et al. (2017).  

The median weight of males at all three sites is, on average, greater than females and hares without a sex identification, though not by much.  The greatest difference can be seen in the Bonanza Riparian site between male weight and NA weight.  There were no juveniles caught at Bonanza Black Spruce that did not have a sex identified. No one site stands out as having the largest/smallest juveniles.


### Juvenile weight comparison (male and female snowshoe hares)

**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for male and female juvenile snowshoe hares. Data: Kielland et al. (2017).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Table  of summary statistics is created.
juv_hare_table <- weights_sex_viz %>% 
  na.omit(sex) %>% 
  group_by(sex) %>% 
  summarize(
    mean_weight = round(mean(weight, na.rm = TRUE),2),
    sd_weight = round(sd(weight, na.rm = TRUE),2),
    n = n())

#kable_styling() is used so that the table is formatted nicely for the knitted document.
juv_hare_table %>% 
  kable(col.names = c("Sex", "Mean weight (g)", "Standard deviation (g)", "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```


```{r, include = FALSE}
#Vectors are created. The pull function will be used to select the sample vectors which will be used in the two sample t-test and effsize test.  NA values are first omitted.
ws_withoutna <- weights_sex_viz %>% 
  na.omit(sex)

female_sample <- ws_withoutna %>% 
  filter(sex == "Female") %>% 
  pull(weight)

male_sample <- ws_withoutna %>% 
  filter(sex == "Male") %>% 
  pull(weight)
```


```{r, include = FALSE}
# Plot histogram and qq plots to check assumptions.
ggplot(data = ws_withoutna, aes(x = weight)) +
  geom_histogram() +
  facet_wrap(~sex) +
  labs(x = "Weight (g)", y = "Count")
# Looks relatively normally distributed for males and females.

ggplot(data = ws_withoutna, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
# Male line more curved.  Female more linear.  Both reasonably linear.
```



```{r, include = FALSE}
# T-test and effect size tests are run.
hare_ttest <- t.test(male_sample, female_sample)
hare_effsize <- cohen.d(male_sample, female_sample)
```

```{r, include = FALSE}
# Mean values are calculated separeately so that they can be used in the in line annotation to explain graphs.
female_mean <- mean(female_sample)
male_mean <- mean(male_sample)
female_sd <- sd(female_sample)
male_sd <- sd(male_sample)

hare_ttest_tidy <- tidy(hare_ttest)
```

On average, juvenile male snowshoe hares have a higher weight than juvenile female hares (`r round(male_mean,2)` $\pm$ `r round(male_sd, 2)` and `r round(female_mean,2)` $\pm$ `r round(female_sd, 2)` g, respectively; mean $\pm$ 1 standard deviation). While the absolute difference in means is `r round(male_mean, 2) - round(female_mean, 2)` g (a `r round((male_mean - female_mean)/((male_mean + female_mean)/2)*100, 2)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(hare_ttest_tidy$parameter,2)`) = `r round(hare_ttest_tidy$statistic,2)`, p < 0.001), and the effect size is small (Cohen's *d* = `r round(hare_effsize$estimate,2)`).



### Relationship between juvenile weight and hind foot length

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Linear model made to view the data
j_hares <- bonanza_hares %>% 
  filter(age == "j")

ggplot(data = j_hares, aes(x = hindft, y = weight)) +geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "gray30") +
  theme_minimal() +
  ggpubr::stat_regline_equation(label.x=60, label.y=1200)+
  labs(x = "Hind Foot Length (mm)",
       y = "Weight (g)")
```

```{r, include = FALSE}
# Do linear regression
juv_hare_lm <- lm(weight ~ hindft, data = j_hares)

# Get tidy versions of the model output to call later in text
juv_hare_lm_tidy <- tidy(juv_hare_lm)
juv_hare_lm_glance <- glance(juv_hare_lm)

# Get Pearson's r correlation 
juv_hare_cor <- cor.test(j_hares$weight, j_hares$hindft)

# Get tidy versions of the model output to call later in text
juv_hare_cor_tidy <- tidy(juv_hare_cor)
```

**Figure 3.** Relationship between hind foot length (mm) and weight (g) for juvenile snowshoe hare. Points indicate individual hare measurements. Linear model summary: $\beta$~1~ = `r round(juv_hare_lm_tidy$estimate[2], 2)` g mm^-1^, p < 0.001, R^2^ = `r round(juv_hare_lm_glance$r.squared,2)`, Pearson's *r* = `r round(juv_hare_cor_tidy$estimate,2)`). 
Data: Kielland et al. (2017).

The relationship between hind foot length and weight does appear linear (Figure 3) This figure was inserted so that the reader could see the spread of the data and visualize the trend. Simple linear regression revealed that the length of the hind foot does significantly predict juvenile hare weight (p < 0.001, R^2^ = `r round(juv_hare_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(juv_hare_lm_tidy$estimate[2], 2)` g mm^-1^ (i.e., for each one millimeter increase in hind foot length we expect an average increase in body weight of `r round(juv_hare_lm_tidy$estimate[2], 2)` g). The Pearson's *r* = `r round(juv_hare_cor_tidy$estimate,2)`, which means `r round(juv_hare_cor_tidy$estimate * 100,2)` percent of variance in the weight is predictable from the hind foot length (a moderately positive correlation).  The p < 0.001.  The data is heteroscedastic.  I am worried that the data does not fit the line extremely closely and that there is a change in variance along the line making the data heteroscedastic. There may be a confounding variable that needs further insight and/or a linear model may not be the best model to predict the trends in the data.  


## Summary

Data analysis reveals the following findings:

- The maximum number of juvenile snowshoe hares were caught in 1999 with the minimum in 2002 and 2009.
- On average, juvenile male snowshoe hares have a higher weight than juvenile female hares (the difference is significant and the effect size is small).
- Weight (g) appears linearly related with hind foot length (mm); the relationship (by simple linear regression) is significant, with a slope of `r round(juv_hare_lm_tidy$estimate[2], 2)` g mm^-1^.
- `r round(juv_hare_cor_tidy$estimate * 100,2)` percent of variance in the weight is predictable from the hind foot length (a moderately positive correlation).

## Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14 


